Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-07-15 pirouette_supplement.tex
Add wall clock times and number of replicates

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{
    Aggregate error distributions, similar to Fig. \ref{fig:example_30} 
    for the main example, but now for a collection of 20 replicate trees. 
    For each setting (true generative, true best candidate, twin generative 
    and twin best candidate), the resulting errors from each replicate 
    pipeline have been merged into a single distribution.
  }
  \label{fig:replicate_trees}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
  \label{fig:example_6_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_32/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
  \label{fig:example_12_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_33/errors.png}
    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
    \label{fig:example_24_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_41/errors.png}
  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
  \label{fig:example_32_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_42/errors.png}
  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
  \label{fig:example_40_taxa}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_19/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
  \label{fig:example_500_nucleotides}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
  \label{fig:example_1000_nucleotides}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_34/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
  \label{fig:example_2000_nucleotides}
\end{figure}

\new{
  From figures \ref{fig:example_500_nucleotides}, 
  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
  we can observe that the discrepancy between the true and twin error 
  distributions tend to become smaller as the number of nucleotides increase. 
  This occurred for both the generative and best candidate cases.
  This follows the expectation that a prior becomes less important when
  more information becomes available.
}
\giovanni{
  Also, do we have the raw data 
  stored somewhere? From what I can judge from the plots the median discrepancy 
  is 0.012 for 500 nucleotides, 0.008 for 1000 nucleotides 
  and 0.003 for 2000 nucleotides.
}
\richel{
  Sure, raw data is at
  \url{https://www.richelbilderbeek.nl/pirouette_example_19.zip},
  \url{https://www.richelbilderbeek.nl/pirouette_example_28.zip}
  and 
  \url{https://www.richelbilderbeek.nl/pirouette_example_34.zip}.
  These URLS can be found on the GitHub site.
  What would you like to do with the raw data?
}
\giovanni{I think it might be nice to have a simple lot of error medians versus number of nucleotides. If we want to be a bit fancier, we can think of including 0.25 and 0.75 percentiles as error bars.}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_22/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
  \label{fig:example_yule}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
  \caption{Aggregate error distributions for 10 replicates. 
  Here each true tree is generated by a BD process. 
  For the inference we used instead a Yule tree prior. This took 
  \richel{unknown} hours to compute.
}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_low.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_mid.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_high.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
  \caption{Aggregate error distributions similar to Fig. \ref{fig:replicate_trees}, but here the number of \new{substitution}s is not imposed to be the same between true and twin alignment. Instead we just use an equal mutation rate. This took 10 hours to compute.}
  \label{fig:example_random_mutations}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_35/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.25 / crown age. This took 11 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_36/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.50 / crown age. This took 14 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_37/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.75 / crown age. This took 16 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_38/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_39/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.50 / crown age. This took 10 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_40/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 2.0 / crown age. This took 19 hours to compute.}
\end{figure}

>>>>

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{
    Aggregate error distributions, similar to Fig. \ref{fig:example_30} 
    for the main example, but now for a collection of 100 replicate trees. 
    For each setting (true generative, true best candidate, twin generative 
    and twin best candidate), the resulting errors from each replicate 
    pipeline have been merged into a single distribution.
    This took 2.7 days (wall clock time) to compute.
  }
  \label{fig:replicate_trees}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each true tree has 6 taxa. 
    This took 2.7 days (wall clock time) to compute.
  }
  \label{fig:example_6_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_32/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each true tree has 12 taxa. 
    This took 6.0 days (wall clock time) to compute.
  }
  \label{fig:example_12_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_33/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each true tree has 24 taxa. 
    This took 9.8 days (wall clock time) to compute.
  }
  \label{fig:example_24_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_41/errors.png}
  \caption{
    Aggregate error distributions for 65 replicates. 
    Here each true tree has 32 taxa. 
    This took 8.0 days (wall clock time) to compute.
  }
  \label{fig:example_32_taxa}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_42/errors.png}
  \caption{
    Aggregate error distributions for 5 replicates. 
    Here each true tree has 40 taxa.
    This took 0.83 days (wall clock time) to compute.
  }
  \label{fig:example_40_taxa}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_19/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each each alignment has a sequence length of 500 nucleotides. 
    This took 2.2 days (wall clock time) to compute.
  }
  \label{fig:example_500_nucleotides}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each each alignment has a sequence length of 1000 nucleotides 
    (as used in the main example). 
    This took 2.7 days (wall clock time) to compute.
  }
  \label{fig:example_1000_nucleotides}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_34/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each each alignment has a sequence length of 2000 nucleotides. 
    This took 4.4 days (wall clock time) to compute.
  }
  \label{fig:example_2000_nucleotides}
\end{figure}

\new{
  From figures \ref{fig:example_500_nucleotides}, 
  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
  we can observe that the discrepancy between the true and twin error 
  distributions tend to become smaller as the number of nucleotides increase. 
  This occurred for both the generative and best candidate cases.
  This follows the expectation that a prior becomes less important when
  more information becomes available.
}
\giovanni{
  Also, do we have the raw data 
  stored somewhere? From what I can judge from the plots the median discrepancy 
  is 0.012 for 500 nucleotides, 0.008 for 1000 nucleotides 
  and 0.003 for 2000 nucleotides.
}
\richel{
  Sure, raw data is at
  \url{https://www.richelbilderbeek.nl/pirouette_example_19.zip},
  \url{https://www.richelbilderbeek.nl/pirouette_example_28.zip}
  and 
  \url{https://www.richelbilderbeek.nl/pirouette_example_34.zip}.
  These URLS can be found on the GitHub site.
  What would you like to do with the raw data?
}
\giovanni{
  I think it might be nice to have a simple lot of error medians 
  versus number of nucleotides. If we want to be a bit fancier, 
  we can think of including 0.25 and 0.75 percentiles as error bars.
}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_22/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each true tree is generated by a Yule process. 
    For the inference we used a Yule tree prior. 
    This took 2.9 days (wall clock time) to compute.
  }
  \label{fig:example_yule}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates. 
    Here each true tree is generated by a BD process. 
    For the inference we used instead a Yule tree prior. 
    This took 2.7 days (wall clock time) to compute.
  }
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_low.png}
  \caption{
    Aggregate error distributions for a distribution of trees, 
    where the true trees are DD with low likelihood.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_mid.png}
  \caption{
    Aggregate error distributions for a distribution of trees, 
    where the true trees are DD with median likelihood.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_high.png}
  \caption{
    Aggregate error distributions for a distribution of trees, 
    where the true trees are DD with high likelihood.
  }
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates.
    similar to Fig. \ref{fig:replicate_trees}, 
    but here the number of \new{substitution}s is not imposed 
    to be the same between true and twin alignment. 
    Instead, an equal mutation rate is used. 
    This took 3.3 days (wall clock time) to compute.
  }
  \label{fig:example_random_mutations}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_35/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 0.25 / crown age. 
    This took 1.8 days (wall clock time) to compute.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_36/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 0.50 / crown age. 
    This took 2.2 days (wall clock time) to compute.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_37/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 0.75 / crown age. 
    This took 2.5 days (wall clock time) to compute.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_38/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 1.25 / crown age. 
    This took 2.9 days (wall clock time) to compute.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_39/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 1.50 / crown age. 
    This took 3.0 days (wall clock time) to compute.
  }
\end{figure}

\begin{figure}[H]
  \includegraphics[width=0.98\textwidth]{pirouette_example_40/errors.png}
  \caption{
    Aggregate error distributions for 100 replicates,
    for the tree distribution presented 
    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
    of 2.0 / crown age. 
    This is done for 100 replicates.
    This took 3.0 days (wall clock time) to compute.
  }
\end{figure}
